Non-parametric approaches to optimal policy are more robust

### 2013-04-08 inst/doc/nonparametric-bayes.Rmd

```{r libraries, include=FALSE}
## Dargh these are ignored if not placed here!?
options(xtable.type = 'latex', table.placement="H", xtable.print.comment=FALSE)
opts_chunk$set(cache.path = 'cache-pdf/', cache=TRUE)
opts_chunk$set(dev = 'Cairo_pdf', fig.width=5.5, fig.height=4)

library(pdgControl)
library(nonparametricbayes)
library(reshape2)
library(ggplot2)
library(data.table)
library(tgp)
library(MCMCpack)
library(plyr)
library(knitcitations)
```

```{r plotting-options, echo=FALSE, include=FALSE}
# This stuff should be handled by knit script(?) (load ggplot2 first though)!
theme_set(theme_bw(base_size=10))
theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
             plot.background = element_rect(fill = "transparent",colour = NA))
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

@@

Most management recommendations from the ecological literature are based
on (or at least motivated by) parametric models. Though in principle
these models reflect mechanistic underpinnings of biological interactions
involved, in practice real mechanisms are often unknown and the true
dynamics too complex to be captured by simple models [@refs]. While
simple mechanistic models can nevertheless provide imporant insights
into possible dynamics -- for instance, demonstrating that a vaccine
does not have to be 100% effective to eliminate the transmission of a
virus  -- such approaches are not well-suited for use in
forecasting outcomes of potential management options.  Non-parametric 
approaches offer a more flexible alternative that can both more accurately
reflect the data available while also representing greater uncertainty
in areas (of state-space) where data is lacking.  Ecological research
and management strategy should pay closer attention to the opportunities 
and challenges nonparametric modeling can offer.

@@

The ecological dynamics of most systems of management interest are
typically both complex and unknown [@refs].  Despite this, quantitative
policy optimization in these systems has almost exclusively been based
on simple, mechanistically motivated parametric models.  There are
several reasons for this.  First, limited available data almost always
percludes any accurate estimate of more complicated models.  Second, the
computational demands of determining an optimal stategy in a sequential
decision-theory problem suffer from the "curse of dimensionality"
that makes it difficult to consider more complicated models, states,
or actions.  Unfortunately, these challenges also make it particularly
difficult to reflect the true uncertainty either in the parameters of
the given models (parametric uncertainty) or in the structure of the
models themselves (structural uncertainty).  Further unknowns such as
measurement uncertainty, parameter uncertainty, unobserved states,
knowledge of boundary conditions, etc. further compound the issue.
Though a hierarchical Bayesian approach provides a natural way to
address these from a statistical standpoint, formulating reasonable
parametric  descriptions of each form of uncertainty is a challenging
task in itself, let alone the computational difficulties of solving such
a system. @Cressie2009 provides a good account of the successess and
challenges of the approach.  Applying these approaches in the management
context of sequential decision making, in which forecasts must be obtained
over a range of possible actions and updated regularly as new information
arrives makes such an approach less feasible still.

@@

<span style="color:#007FFF;">***Okay, perhaps that's novel, but it's pretty trivial.  Isn't it
obvious to everyon that it's trivial?***</span>

@@

The essense of the GP approach can be captured in the
following thought experiment: An exhaustive parametric approach to the
challenge of structural uncertainty might proceed by writing down all
possible functional forms for the underlying dynamical system with all
possible parameter values for each form, and then consider searching
over this huge space to select the most likely model and parameters;
or using a Bayesian approach, assign priors to each of these possible
models and infer the posterior distribution of possible models. The
GP approach can be thought of as a computationally
efficient approximation to this approach. GPs represent
a large class of models that can be though of as capturing or reasonably
approximating the set of models in this collection.  By modeling at the
level of the process, rather than the level of parametric equation,
we can more concisely capture the possible behavior of these curves.
In place of a parametric model of the dynamical system, the GP 
approach postulates a prior distribution of (n-dimensional)
curves that can be though of as approximations to a range of possible
(parametric) models that might describe the data. The GP allows us
to consider a set of possible curves simultaneously.  
<span style="color:#007FFF;"> ***Figure 1 include curves drawn from the posterior density?*** </span>

@@

Once the posterior Gaussian process (GP) has been estimated [e.g. see
@Munch2005], it is necessary to adapt it in place of the parametric
equation for the stochastic dynamic programming (SDP) solution [see
@Mangel1988 for a detailed description of parametric SDP methods] to the
optimal policy. The essense of the idea is straight forward -- we will use
the estimated GP in place of the parametric growth function to determine
the stochastic transition matrix on which the SDP calculations are based.
The SDP is solved in a discritized state space -- both the continuously 
valued population densities $X$ and harvest quotas $h$ are first mapped to
a bounded, discrete grid.  (For simplicity we will consider a uniform grid,
though for either parametric or GP-based SDP it is often advantageous to 
use a non-uniform discretization such as a basis function representation,
e.g. see [@Deisenroth2009]).

The SDP approach then computes a transition matrix, $\mathbf{F}$.  We 
demonstrate that calculation is just as straight forward based on the GP
as it is in the classical context using the parametric model.  The 
${i,j}$ of the transition matrix $F$ entry gives the probability of transitioning into 
state $x_i$ given that the system is in state $x_j$ in the previous 
timestep.  To generate the transition
matrix based on the posterior GP, we need only the expected values
at each grid point and the corresponding variances (the diagonal of
the covariance matrix), as shown in Figure 1.  Given the mean of the 
GP posterior at each gridpoint as the vector $E$ and variance at that
point as vector $V$, the probability of transitioning from state $x_i$ to state $x_j$ is

@@

**Pseudocode for the determining the transtion matrix from the GP**

@@

Using the discrete transition matrix we may write down the Bellman
recursion defining the the stochastic dynamic programming iteration:

@@

where $V(x_t)$ is the value of being at state $x$ at time $t$, $h$
is control (harvest level) chosen. Numerically, the maximization is
accomplished as follows. Consider the set of possible control values to
be the discrete values corresponding the the grid of
stock sizes.  Then for each $h_t$ there is a corresponding transition
matrix $\mathbf{F}_h$ determined as described above but with mean 
$\mu = x_j - h_t$. Let $\vec{V_t}$ be the vector whose $i$th element corresponds
to the value of having stock $x_i$ at time $t$.  Then let $\Pi_h$ be
the vector whose $i$th element indicates the profit from harvesting
at intensity $h_t$ given a population $x_i$ (e.g. $\max(x_i, h_t)$
since one cannot harvest more fish then the current population size).
Then the Bellman recursion can be given in matrix form as

@@

Concerns over the potential for tipping points in ecological dynamics
[@Scheffer2001] highlight the dangers of uncertainty in ecological
management and pose a substantial challenge to existing decision-theoretic
approaches [@Brozovic2011].  Because intervention is often too late 
after a tipping point has been crossed (but see @Hughes2013), management
is most often concerned with avoiding potentially catostrophic tipping
points before any data is available at or following a transition that
would more clearly reveal these regime shift dyanmics [e.g. @Bestelmeyer2012].

@@

The maximum likelihood estimates (MLE) will assume that the process noise $Z_t$
is log-normal, as in the simulation itself.  In reality the noise structure might 
differ from this assumption, making this a best-case scenario for the MLE approach.  
The GP must also make an assumption about the noise structure -- in this case we 
assume additve noise, though multiplicative noise would be more accurate.  This
will illustrate the flexibility of the GP in fitting different noise structures
even without knowledge of the correct assumptions.

The GP estimate is Bayesian approach which requires specification of
prior distributions for the hyperparameters [see Appendix for details,
or @Rassmussen1996, @Munch2005 for background], as given in Table 1.
Then given the data, both the GP and the MLE parametric models can
be estimated as described above, as we implement in R in the code
provided in supplementary materials.

@@

Figure 1 shows the  inferred Gaussian Process compared to the true and the MLE parametric
models. The grey shading shows the uncertainty estimated by the GP.  Unlike the MLE estimates,
the GP uncertainty changes depending on where the data is available, allowing for a much 
greater level of uncertainty at low stock sizes.  While it would be natural (and straight forward)
to condition the GP on passing through the origin (0,0) (see appendix), the estimate shown
here is based only on the observed data.

@@

## Figure 2:

_The take-home message, showing that the GP is closest to the optimal
strategy, while the parametric methods are less accurate.  Visualizing the
policy may be more useful for the technical reader, the general audience
may prefer Figure 3 showing all replicates of the population collapse
under the parametric model and not under the GP._

```{r policies_plot, dependson=c("opt", "plotting-options"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.  Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979]."}
policies <- 
  melt(data.frame(stock=x_grid, 
                  GP = x_grid[OPT$gp_D], 
                  Parametric = x_grid[OPT$est_D],
                  True = x_grid[OPT$true_D],
                  Structural = x_grid[OPT$alt_D]),
       id="stock")
names(policies) <- c("stock", "method", "value")

@@

## Figure 3:

_Figure 3 is a less abstract and more visceral visualization of the
take-home message, with the structurally inaccurate model leading
universally to a collapse of the fishery and very few profits, while the
Gaussian process performs nearly optimally.  The parametric approach even
with the correct underlying structure does not perform optimally, choosing
in this case to under-fish (may need to show harvest dynamics since
that is not clear from the figure! Also isn't general, sometimes does
optimally, sometimes over-fishes.  Perhaps need to show more examples.)
May need to show profits too?_

```{r sim_plot, dependson=c("sim", "plotting-options"), fig.cap="Gaussian process inference outperforms parametric estimates. Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model."}
ggplot(dt) + 
    geom_line(aes(time, fishstock, 
                  group=interaction(reps, method), 
                  color=method), alpha=.1) +
    scale_colour_manual(values=cbPalette, 
                        guide = guide_legend(override.aes = list(alpha = 1)))
```

## Figure 4:

_Shows the sensitivity analysis.  A histogram of distribution of yield
over stochastic realizations, showing that the qualitative results do
not depend on the stochastic realization of the training data here, or on
the parameters of the underlying model, though quantitative differences
are visible._

### Discussion on how we compare performance of policies

* Replicate stochastic simulations 
* Sensitivity analysis (Figure 4).

Discussion / Conclusion
==================================================================

Non-parametric methods have received far too little attention in
ecological modeling efforts that are aimed at improved conservation
planning and decision making support.

* Adapting a non-parametric approach requires modification of existing
methods for decision theory.  We have illustrated how this might be
done in the context of stochastic dynamic programming, opening the
door for substantial further research into how these applications might
be improved.

Three elements contribute to the machine learning model performing as well
or better than the mechanisitic models in this context.

### 1. Relevant state space

The dynamics occur over a range of state-space in which the machine learning
model closely matches the predictions of y
implausible model performs as well or better than the more plausible
model.

This is the most obvious and immediate reason why the shortcomings
that appear to make the model implausible do not make it useless.
This effect is enhanced in our example because the management actions help
drive the system towards rather than away from this region of state-space.

### 2. Predictive accuracy

This application relies only on the predictive accuracy of the model,
not an interpretation of the parameters.

<span style="color:#007FFF;"> wow, run-on tangent? </span>
Predictive accuracy is not the goal of all modeling, as ecologists
have been observing for as long as they made models (perhaps none
more memorably than @Levins1969).  Mechanistic modeling is at its
most powerful not when it is used to directly forecast future states
but when it provides an understanding of how to approach a problem.
SIR-type models from the epidemiological literature are a good example.
While the simplest SIR models have little predictive power over the
outbreak intensity or timing at a particular location, they provide a
powerful understanding of the spread of an infection in terms of a single,
biologically meaningful parameter: $R_0$, the basic reproductive number.
From the model, it becomes clear that management need not vaccinate every
member of the population to stop the spread, but rather it suffices to
vaccinate a sufficient fraction of the population to reduce $R_0$ below 1.

@@

2. Better express ambiguity in regions where data is not available

One of the greatest strengths of mechanistic models is their greatest weakness as well.

@@

<span style="color:#007FFF;"> Jargony and vague </span>
While we have highlighted certain generic of this problem that allow the
nonparametric approach to excel -- short timescales between successive
observations and actions, the accuracy in the appropriate region of state
space, the ability to express uncertainty outside the observed data --
there are equally several aspects for which the nonparametric approach is
at a relative disadvantage.

<span style="color:#007FFF;"> Model complexity </span>
In this simulated example, the underlying
dynamics are truly governed by a simple parametric model, allowing
the parametric approaches to be more accurate.  Similarly, because the
dynamics are  one-dimensional dynamics and lead to  stable nodes (rather
than other attractors such as limit-cycles resulting in oscillations),
the training data provides relatively limited information about the
dynamics.  For these reasons, we anticipate that in higher-dimensional
examples characteristic of ecosystem management problems that the machine
learning approach will prove even more valuable.

<span style="color:#007FFF;"> Data complexity. Perhaps too far out of scope... </span>
The machine learning approach is also better suited to complex and
disparate data.  Incorporating various sources of information into
mechanistic models can be an immensely difficult due to the increased
complexity involved.  Only thanks to tandem advances in increasing
computational power and hierarchical statistical methodology have we
been able to tackle such intuitively important complexity (and the
potentially new available data that accompanies it) such as spatial
distribution, heterogeneities of space, time, and individuals, to shift
to ecosystem-based approaches from single-species based approaches.
Without the need to formulate mechanisms, many modern machine learning
algorithms can leverage potential information from all available sources
of data directly.  The algorithms can recognize unanticipated or subtle
patterns in large data sets that enable more accurate predictions than
mechanistic models that are formulated at a more macroscopic level.

<span style="color:#007FFF;"> 
* Discuss constant escapement in model, in policies.

* Limitations of this comparison: Are the maximum-likelihood solutions
a straw man?

* Discussion of alternative related approaches: POMDP/MOMDP
</span>

* Multiple species 
* Online learning 
* Multiple step-ahead predictions 
* Explicitly accomidating additional uncertainties 
* Improving inference of optimal policy from the GP

```{r echo=FALSE, results="asis"}
#bibliography()
```

@@

@@

### The Guassian process

@@

The posterior Gaussian process is completely defined by the expected value
and covariance matrix at a defined set of training points.  For simplicty
we will consider a these points to fall on a discrete, uniform grid $x$
of `r length(x_grid)` points from `r min(x_grid)` to `r max(x_grid)`
(1.5 times the positive equilibrium $K$).  Again to keep things simple
we will use this same grid discritization for the parametric approach.
Other options for choosing the grid points, including collocation methods
and functional basis expansion (or even using Guassian processes in place
of the discrete optimization; an entirely different context in which GP
can be used in SDP, see [@Deisenroth2009]) could also be considered.

The transition matrix $\mathbf{F}$ is thus an `r length(x_grid)`
by `r length(x_grid)` matrix for which the ${i,j}$ entry gives the
probability of transitioning into state $x_i$ given that the system is
in state $x_j$ in the previous timestep.  To generate the transition
matrix based on the posterior GP, we need only the expected values
at each grid point and the corresponding variances (the diagonal of
the covariance matrix), as shown in Figure 1.  Given the mean at each
gridpoint as the length `r length(x_grid)` vector $E$ and variance $V$,
the probability of transitioning from state $x_i$ to state $x_j$ is
simply $\mathcal{N}\left(x_j | \mu = E_i, \sigma = \sqrt{V_i}\right)$,
where $\mathcal{N}$ is the Normal density at $x_j$ with mean $\mu$ and
variance $\sigma^2$.  Strictly speaking, the transition probability should
be calculated by integrating the normal density over the bin of width
$\Delta$ centered at $x_j$.  For a sufficiently fine grid that $f(x_j)
\approx f(x_j + \Delta)$, it is sufficient to calculate the density at
$x_j$ and then row-normalize the transition matrix.

@@

>>>>

```{r libraries, include=FALSE, cache=FALSE}
library(pdgControl)
library(nonparametricbayes)
library(reshape2)
library(ggplot2)
library(data.table)
library(tgp)
library(MCMCpack)
library(plyr)
library(knitcitations)
```

```{r plotting-options, echo=FALSE, include=FALSE, cache=TRUE}
## Dargh these are ignored if not placed here!?
options(xtable.type = 'latex', table.placement="H", xtable.print.comment=FALSE)
opts_chunk$set(cache.path = 'cache-pdf/', cache=TRUE)
opts_chunk$set(dev = 'Cairo_pdf', fig.width=5.5, fig.height=4)

# This stuff should be handled by knit script(?) (load ggplot2 first though)!
theme_set(theme_bw(base_size=10))
theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
             plot.background = element_rect(fill = "transparent",colour = NA))
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

@@

Most management recommendations from the ecological literature are based
on (or at least motivated by) parametric models. Though in principle
these models reflect mechanistic underpinnings of biological interactions
involved, in practice real mechanisms are often unknown and the true
dynamics too complex to be captured by simple models [@refs]. While
simple mechanistic models can nevertheless provide important insights
into possible dynamics -- for instance, demonstrating that a vaccine
does not have to be 100% effective to eliminate the transmission of a
virus  -- such approaches are not well-suited for use in
forecasting outcomes of potential management options.  Non-parametric 
approaches offer a more flexible alternative that can both more accurately
reflect the data available while also representing greater uncertainty
in areas (of state-space) where data is lacking.  Ecological research
and management strategy should pay closer attention to the opportunities 
and challenges nonparametric modeling can offer.

@@

The ecological dynamics of most systems of management interest are
typically both complex and unknown [@refs].  Despite this, quantitative
policy optimization in these systems has almost exclusively been based
on simple, mechanistically motivated parametric models.  There are
several reasons for this.  First, limited available data almost always
precludes any accurate estimate of more complicated models.  Second, the
computational demands of determining an optimal strategy in a sequential
decision-theory problem suffer from the "curse of dimensionality"
that makes it difficult to consider more complicated models, states,
or actions.  Unfortunately, these challenges also make it particularly
difficult to reflect the true uncertainty either in the parameters of
the given models (parametric uncertainty) or in the structure of the
models themselves (structural uncertainty).  Further unknowns such as
measurement uncertainty, parameter uncertainty, unobserved states,
knowledge of boundary conditions, etc. further compound the issue.
Though a hierarchical Bayesian approach provides a natural way to
address these from a statistical standpoint, formulating reasonable
parametric  descriptions of each form of uncertainty is a challenging
task in itself, let alone the computational difficulties of solving such
a system. @Cressie2009 provides a good account of the successes and
challenges of the approach.  Applying these approaches in the management
context of sequential decision making, in which forecasts must be obtained
over a range of possible actions and updated regularly as new information
arrives makes such an approach less feasible still.

@@

<span style="color:#007FFF;">***Okay, perhaps that's novel, but it's pretty trivial.  Isn't it
obvious to everyone that it's trivial?***</span>

@@

The essence of the GP approach can be captured in the
following thought experiment: An exhaustive parametric approach to the
challenge of structural uncertainty might proceed by writing down all
possible functional forms for the underlying dynamical system with all
possible parameter values for each form, and then consider searching
over this huge space to select the most likely model and parameters;
or using a Bayesian approach, assign priors to each of these possible
models and infer the posterior distribution of possible models. The
GP approach can be thought of as a computationally
efficient approximation to this approach. GPs represent
a large class of models that can be though of as capturing or reasonably
approximating the set of models in this collection.  By modeling at the
level of the process, rather than the level of parametric equation,
we can more concisely capture the possible behavior of these curves.
In place of a parametric model of the dynamical system, the GP 
approach postulates a prior distribution of (n-dimensional)
curves that can be though of as approximations to a range of possible
(parametric) models that might describe the data. The GP allows us
to consider a set of possible curves simultaneously.  
<span style="color:#007FFF;"> ***Figure 1 include curves drawn from the posterior density?*** </span>

@@

Once the posterior Gaussian process (GP) has been estimated [e.g. see
@Munch2005], it is necessary to adapt it in place of the parametric
equation for the stochastic dynamic programming (SDP) solution [see
@Mangel1988 for a detailed description of parametric SDP methods] to the
optimal policy. The essence of the idea is straight forward -- we will use
the estimated GP in place of the parametric growth function to determine
the stochastic transition matrix on which the SDP calculations are based.
The SDP is solved in a discretized state space -- both the continuously 
valued population densities $X$ and harvest quotas $h$ are first mapped to
a bounded, discrete grid.  (For simplicity we will consider a uniform grid,
though for either parametric or GP-based SDP it is often advantageous to 
use a non-uniform discretization such as a basis function representation,
e.g. see [@Deisenroth2009]).

The SDP approach then computes a transition matrix, $\mathbf{F}$.  We 
demonstrate that calculation is just as straight forward based on the GP
as it is in the classical context using the parametric model.  The 
${i,j}$ of the transition matrix $F$ entry gives the probability of transitioning into 
state $x_i$ given that the system is in state $x_j$ in the previous 
time-step.  To generate the transition
matrix based on the posterior GP, we need only the expected values
at each grid point and the corresponding variances (the diagonal of
the covariance matrix), as shown in Figure 1.  Given the mean of the 
GP posterior at each grid-point as the vector $E$ and variance at that
point as vector $V$, the probability of transitioning from state $x_i$ to state $x_j$ is

@@

**Pseudocode for the determining the transition matrix from the GP**

@@

Using the discrete transition matrix we may write down the Bellman
recursion defining the stochastic dynamic programming iteration:

@@

where $V(x_t)$ is the value of being at state $x$ at time $t$, $h$
is control (harvest level) chosen. Numerically, the maximization is
accomplished as follows. Consider the set of possible control values to
be the discrete values corresponding the grid of
stock sizes.  Then for each $h_t$ there is a corresponding transition
matrix $\mathbf{F}_h$ determined as described above but with mean 
$\mu = x_j - h_t$. Let $\vec{V_t}$ be the vector whose $i$th element corresponds
to the value of having stock $x_i$ at time $t$.  Then let $\Pi_h$ be
the vector whose $i$th element indicates the profit from harvesting
at intensity $h_t$ given a population $x_i$ (e.g. $\max(x_i, h_t)$
since one cannot harvest more fish then the current population size).
Then the Bellman recursion can be given in matrix form as

@@

Concerns over the potential for tipping points in ecological dynamics
[@Scheffer2001] highlight the dangers of uncertainty in ecological
management and pose a substantial challenge to existing decision-theoretic
approaches [@Brozovic2011].  Because intervention is often too late 
after a tipping point has been crossed (but see @Hughes2013), management
is most often concerned with avoiding potentially catostrophic tipping
points before any data is available at or following a transition that
would more clearly reveal these regime shift dynamics [e.g. @Bestelmeyer2012].

@@

The maximum likelihood estimates (MLE) will assume that the process noise $Z_t$
is log-normal, as in the simulation itself.  In reality the noise structure might 
differ from this assumption, making this a best-case scenario for the MLE approach.  
The GP must also make an assumption about the noise structure -- in this case we 
assume additive noise, though multiplicative noise would be more accurate.  This
will illustrate the flexibility of the GP in fitting different noise structures
even without knowledge of the correct assumptions.

The GP estimate is Bayesian approach which requires specification of
prior distributions for the hyper-parameters [see Appendix for details,
or @Rassmussen1996, @Munch2005 for background], as given in Table 1.
Then given the data, both the GP and the MLE parametric models can
be estimated as described above, as we implement in R in the code
provided in supplementary materials.

@@

Figure 1 shows the  inferred Gaussian Process compared to the true and the
MLE parametric models, predicting the relationship between observed stock
size (x-axis) to the stock size following recruitment the following year.
The gray shading shows the uncertainty estimated by the GP.  Unlike the
MLE estimates, the GP uncertainty changes depending on where the data is
available, allowing for a much greater level of uncertainty at low stock
sizes.  While it would be natural (and straight forward) to condition
the GP on passing through the origin (0,0) (see appendix), the estimate
shown here is based only on the observed data. The observed data from
which each model is estimated is also shown.  The observations come
from only a limited region of state space corresponding to unharvested
or weakly harvested system.  No observations occur at the theoretical
optimum (7.5 units, see Figure 2) or near the tipping point (5 units).

@@

The resulting optimal management strategy based on each of the inferred
models is shown in Figure 2, against the optimal strategy given the
true underlying dynamics.  Policies are shown in terms of target
escapement, $S_t$.  Under models such as this a constant escapement
policy is expected to be optimal [@Reed1979], whereby population levels
below a certain size $S$ are unharvested, while above that size the harvest
strategy aims to return the population to $S$, resulting in the hockey-stick
shaped policies shown.  Only the GP policy deviates slightly from this.  The MLE parametric model
corresponding to parametric uncertainty alone (gold line labeled "Parametric", 
that is, using the same structural form as the underlying simulation) nevertheless overestimates 
the carrying capacity of the stock and therefore underestimates the harvest
rates, always waiting for the stock to rise further than is actually possible.
The Ricker model (green line, showing "Structural" inaccuracies) does not 
foresee the tipping point dynamics reflected in the concavity of the other three
models (Figure 1), and selects a much more aggressive harvesting strategy.

```{r policies_plot, dependson=c("opt", "plotting-options"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.  "}
policies <- 
  melt(data.frame(stock=x_grid, 
                  GP = x_grid[OPT$gp_D], 
                  Parametric = x_grid[OPT$est_D],
                  True = x_grid[OPT$true_D],
                  Structural = x_grid[OPT$alt_D]),
       id="stock")
names(policies) <- c("stock", "method", "value")

@@

The consequences of managing 100 replicate realizations of the 
simulated fishery under each of the policies estimated is shown in Figure 3.  As expected
from the policy curves, the structurally correct model ("Parametric" uncertainty, gold) 
underharvests, leaving the stock to vary around it's unfished optimum.  The GP overharvests
the population slightly relative to the optimal value.  The structurally
incorrect Ricker model ("Structural", green line) overharvests the population passed
the tipping point consistently, resulting in the immediate crash of the stock and 
thus derives minimal profits.

***(may need to show harvest dynamics since
that is not clear from the figure! Also isn't general, sometimes does
optimally, sometimes over-fishes.  Perhaps need to show more examples.)
May need to show profits too?***

```{r sim_plot, dependson=c("sim", "plotting-options"), fig.cap="Gaussian process inference outperforms parametric estimates. Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model."}
ggplot(dt) + 
    geom_line(aes(time, fishstock, 
                  group=interaction(reps, method), 
                  color=method), alpha=.1) +
    scale_colour_manual(values=c("GP" = cbPalette[1], "Parametric" = cbPalette[2], "True" = cbPalette[3], "Structural" = cbPalette[4]),  
                        guide = guide_legend(override.aes = list(alpha = 1)))
```

The results shown in Figures 1-3 are not unique to the simulated data or models chosen
here, but arises across a range of parameter values and simulations as shown in the 
supplemental figures.  The results across this range can most easily be compared 
by the relative differences in net present value realized by each of the approaches,
as shown in Figure 4.  The Gaussian Process most consistently realizes a value 
close to the optimal solution, and importantly avoids ever driving the system across
the tipping point, which results in the near-zero value cases in the parametric models.

![Histograms of the realized net present value of the fishery over a range of simulated data and resulting parameter estimates. For each data set, the three models are estimated as described above. Values plotted are the averages of a given policy over 100 replicate simulations. Details and code provided in the supplement.](figure/figure4.png)

Discussion 
==================================================================

Non-parametric Bayesian methods have received far too little attention
in ecological modeling efforts that are aimed at improved conservation
planning and decision making support.  Such approaches may be particularly
useful when the available data is restricted to a limited area of
state-space, which can lead parametric models to underestimate the
uncertainty in dynamics at population levels (states) which have not
been observed.  One reason for the relative absence of nonparametric
approaches in the natural resource management context may be the lack
of existing approaches for adapting the non-parametric Bayesian models
previously proposed [@Munch2005] to a decision-theoretic framework.
Adapting a non-parametric approach requires modification of existing
methods for decision theory.  We have illustrated how this might be
done for a classic stochastic dynamic programming problem, opening the
door for substantial further research into how these applications might
be improved.

While non-parametric Bayesian approaches will not always be preferable 
to simple mechanistic models, we highlight three aspects of the problem
consider here that make these methods particularly valuable.  These aspects
are common to many conservation decision making problems, which thus merit
greater use of non-parametric machine learning approaches that can best 
take advantage of them.

### 1. Large uncertainty where the data is poor

The parametric models perform worst when they propose a management strategy
outside the range of the observed data. The non-parametric Bayesian approach, 
in contrast, allows a predictive model that expresses a great deal of uncertainty
about the probable dynamics outside the observed range, while retaining very
good predictive accuracy in the range observed.  The management policy 
dictated by the GP balance this uncertainty against
the immediate value of the harvest, and act to stabilize the population 
dynamics in a region of state space in which the predictions can be 
reliably reflected by the data.

### 2. Predictive accuracy where data is good

While expressing larger uncertainty outside the observed data, the GP
can also provide a better fit with smaller uncertainty inside the range
of the observed data. This arises from the greater flexibility of the 
Gaussian process, which describes a large family of possible curves.  
While in a parametric context this over-fitting would be more worrisome --
a high-degree polynomial could fit the data even better -- those 
concerns are driven by the resulting parametric fit outside the data,
which may involve wild oscillations unsupported by the data.  As we 
have seen in #1, the GP is less vulnerable to such unjustified predictions
outside the data, and is meanwhile free to benefit from the greater 
fit where the data is available.

<!--
<span style="color:#007FFF;"> ***wow, run-on tangent here?*** </span>
Predictive accuracy is not the goal of all modeling, as ecologists
have been observing for as long as they made models (perhaps none
more memorably than @Levins1969).  Mechanistic modeling is at its
most powerful not when it is used to directly forecast future states
but when it provides an understanding of how to approach a problem.
SIR-type models from the epidemiological literature are a good example.
While the simplest SIR models have little predictive power over the
outbreak intensity or timing at a particular location, they provide a
powerful understanding of the spread of an infection in terms of a single,
biologically meaningful parameter: $R_0$, the basic reproductive number.
From the model, it becomes clear that management need not vaccinate every
member of the population to stop the spread, but rather it suffices to
vaccinate a sufficient fraction of the population to reduce $R_0$ below 1.
-->

@@

@@

<span style="color:#007FFF;"> Jargony and vague </span>
While we have highlighted certain generic of this problem that allow the
nonparametric approach to excel -- short timescales between successive
observations and actions, the accuracy in the appropriate region of state
space, the ability to express uncertainty outside the observed data --
there are equally several aspects for which the nonparametric Bayesian approach is
nevertheless at a relative disadvantage. In this simulated example, the underlying
dynamics are truly governed by a simple parametric model, allowing
the parametric approaches to be more accurate.  Similarly, because the
dynamics are  one-dimensional dynamics and lead to  stable nodes (rather
than other attractors such as limit-cycles resulting in oscillations),
the training data provides relatively limited information about the
dynamics.  For these reasons, we anticipate that in higher-dimensional
examples characteristic of ecosystem management problems that the machine
learning approach will prove even more valuable.

<span style="color:#007FFF;"> Data complexity. Perhaps too far out of scope... </span>
The nonparametric Bayesian approach is also better suited to complex and
disparate data.  Incorporating various sources of information into
mechanistic models can be an immensely difficult due to the increased
complexity involved.  Only thanks to tandem advances in increasing
computational power and hierarchical statistical methodology have we
been able to tackle such intuitively important complexity (and the
potentially new available data that accompanies it) such as spatial
distribution, heterogeneities of space, time, and individuals, to shift
to ecosystem-based approaches from single-species based approaches.
Without the need to formulate mechanisms, many modern machine learning
algorithms can leverage potential information from all available sources
of data directly.  The algorithms can recognize unanticipated or subtle
patterns in large data sets that enable more accurate predictions than
mechanistic models that are formulated at a more macroscopic level.

@@

<span style="color:#007FFF;"> 
***Possible additional discussion material?***

* Discuss constant escapement in model, in policies.
* Limitations of this comparison: Are the maximum-likelihood solutions
a straw man?
* Discussion of alternative related approaches: POMDP/MOMDP

* Online learning 
* Multiple step-ahead predictions 
* Explicitly accomidating additional uncertainties 
* Improving inference of optimal policy from the GP
</span>

@@

### The Gaussian process

@@

The posterior Gaussian process is completely defined by the expected value
and covariance matrix at a defined set of training points.  For simplicity
we will consider a these points to fall on a discrete, uniform grid $x$
of `r length(x_grid)` points from `r min(x_grid)` to `r max(x_grid)`
(1.5 times the positive equilibrium $K$).  Again to keep things simple
we will use this same grid discretization for the parametric approach.
Other options for choosing the grid points, including collocation methods
and functional basis expansion (or even using Gaussian processes in place
of the discrete optimization; an entirely different context in which GP
can be used in SDP, see [@Deisenroth2009]) could also be considered.

@@

***Can we get this figure into the main text with appropriate discussion without getting too technical?***
